---
title: "CDC National Health and Nutrition Survey Data Analysis"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
require("RCurl")
require("dplyr")
require("plyr")
require("tidyr")
require("ggplot2")
require("ggthemes")
library(shiny)
require(car)
library(plotly)
```
-----------------------------------------------------------------------
### Bar Graph: Depression v. Annual Income by Marital Status
```{r, echo=FALSE}
selectInput(input = "marital",
                     label = "Select a Martial Status: ",
                     c("Divorced", "Living with partner", "Married", "Never married", "Separated", "Widowed"), selected = "Married")
plotlyOutput("barplot",width = "auto", height = "auto")
  output$barplot<-renderPlotly({
  df1<- read.csv("mentalHealth.reformatted.csv", stringsAsFactors = FALSE)
  df2<- read.csv("demographics.reformatted.csv", stringsAsFactors = FALSE)
  df<-merge(df1, df2)
    df <-df %>%
      dplyr::select(MARITAL, ANNUAL_HH_INCOME, DEPRESSED)%>%
      dplyr::filter(MARITAL%in%(input$marital), !ANNUAL_HH_INCOME%in%c("null", "Under $20,000", "$20,000 + "), DEPRESSED%in%c("Never", "More than half the days", "Nearly every day", "Several days"))%>%
      dplyr::group_by(MARITAL, ANNUAL_HH_INCOME, DEPRESSED)%>%
      dplyr::summarize(count = n())%>%dplyr::mutate(percentage = count/sum(count), pos = (cumsum(percentage) - 0.5 * percentage))
    #mapping strings to factors for income brackets for ordering
    df$scode<-mapvalues(df$ANNUAL_HH_INCOME, from = c("$0 - 4,999", "$5,000 - 9,999", "$10,000 - 14,999", "$15,000 - 19,999", "$20,000 - 24,999", "$25,000 - 34,999", "$35,000 - 44,999", "$45,000 - 54,999", "$55,000 - 64,999", "$65,000 - 74,999", "$75,000 - 99,999", "$100,000 +"), to=c("1", "2", "3", "4", "5", "6","7", "8", "9","10", "11", "12"))
    
      p<-ggplot(data= df, aes(x =ANNUAL_HH_INCOME, y=percentage, fill=DEPRESSED)) + geom_bar(stat="identity", position ="stack") + labs(title = "") + ggtitle("") + theme(legend.title =element_text(size=8), legend.text = element_text(size=6), plot.title=element_text(size=10, face="bold"), axis.title.x = element_text(size=7), axis.text=element_text(size=6)) + scale_x_discrete(limits= c("$0 - 4,999", "$5,000 - 9,999", "$10,000 - 14,999", "$15,000 - 19,999", "$20,000 - 24,999", "$25,000 - 34,999", "$35,000 - 44,999", "$45,000 - 54,999", "$55,000 - 64,999", "$65,000 - 74,999", "$75,000 - 99,999", "$100,000 +"), expand=c(0,-2)) + xlab("") + ylab("% of Total Count Depressed") + geom_text(data = subset(df,percentage>=0.05), aes(y=pos, label = paste(round(percentage*100,1), "%")), size=3) + scale_y_continuous(expand=c(0,0)) + coord_flip()
ggplotly(p)%>%layout(margin=list(b=150))
})
```

***
The happiest divorced individuals are in $75,000 - 99,000 income bracket with a never depressed rate of 91.7%. 

90% of the respondents in $65,000 to 74,999 income bracket who are also living with a partner are never depressed. 

Widowers in the $100,000 + bracket are the happiest with a never depressed rate of 80%. 

Separated individuals in the $100,000 + bracket are the happiest with 100% of subjects never being depressed. 

Single people are relatively happy across all income brackets, but the highest never depressed rate is in the $100,000 + bracket at 85.9%.

The happiest married people are in the highest income bracket of $100,000 + with a never depressed rate of 87.7%. 

In general, rates of depression are lower for respondents who are also high-income earners across all marriage categories. 

### Box Plot: Ratio Income Poverty v. Suicidal Thoughts by Gender
```{r, echo=FALSE}
selectInput(input = "gender",
              label = "Select a Gender: ",
              c("Male", "Female"), selected = "Male")
plotlyOutput("boxplot", width = "auto", height = "auto")
output$boxplot<-renderPlotly({
  df1<- read.csv("mentalHealth.reformatted.csv", stringsAsFactors = FALSE)
  df2<- read.csv("demographics.reformatted.csv", stringsAsFactors = FALSE)
  df<-merge(df1, df2)
  df$scode<-mapvalues(df$BETTER_DEAD, from = c("Nearly every day", "More than half the days", "Several days", "Never"), to=c("1", "2", "3", "4"))
      df <-df %>%
      dplyr::select(GENDER, RATIO_INCOME_POVERTY, BETTER_DEAD)%>%
      dplyr::filter(GENDER%in%c(input$gender), RATIO_INCOME_POVERTY!="null",BETTER_DEAD%in%c("More than half the days", "Nearly every day", "Several days", "Never"))%>%
      dplyr::group_by(GENDER, RATIO_INCOME_POVERTY, BETTER_DEAD)
      p<-ggplot(data= df, aes(x =BETTER_DEAD, y=as.numeric(as.character(RATIO_INCOME_POVERTY)), fill=BETTER_DEAD)) + geom_boxplot() + xlim(c("Nearly every day", "More than half the days", "Several days", "Never")) + ylim(0,5) + xlab("Feels Suicidal") + ylab("Ratio Income Poverty") + theme(legend.position="none") 
ggplotly(p)%>%layout(margin=list(b=150))
})
    

```

***
For male respondents in this survey, those who feel suicidal nearly everyday have a median income to poverty ratio of 0.63 while those who are never suicidal have a median ratio of 2.19. In general, for male respondents, as the severity of depression increases the income to poverty ratio decreases.

Female respondents who are suicidal nearly every day and more than half the days have a median income to poverty ratio of 1 and 0.91 respectively. Females who feel suicidal for several days have a median income to poverty ratio of 1.30. Notice that females who are never suicidal have a median income to poverity ratio of 2.02.
