---
title: "CDC National Health and Nutrition Survey Data Analysis"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
require("jsonlite")
require("RCurl")
require("dplyr")
require("plyr")
require("tidyr")
require("ggplot2")
require("ggthemes")
library(shiny)
require(car)
library(plotly)
require(rbokeh)
```
-----------------------------------------------------------------------
### Bar Graph: Depression v. Annual Income by Marital Status
```{r, echo=FALSE}
#Here is the server script
selectInput(input = "marital",
                     label = "Select a Martial Status: ",
                     c("Divorced", "Living with partner", "Married", "Never married", "Separated", "Widowed"), selected = "Married")
plotlyOutput("barplot",width = "auto", height = "auto")
  output$barplot<-renderPlotly({
df1<- read.csv("mentalHealth.reformatted.csv", stringsAsFactors = FALSE)
df2<- read.csv("demographics.reformatted.csv", stringsAsFactors = FALSE)
df<-merge(df1, df2)
    df <-df %>%
      dplyr::select(MARITAL, ANNUAL_HH_INCOME, DEPRESSED)%>%
      dplyr::filter(MARITAL%in%(input$marital), !ANNUAL_HH_INCOME%in%c("null", "Under $20,000", "$20,000 + "), DEPRESSED!="null" )%>%
      dplyr::group_by(MARITAL, ANNUAL_HH_INCOME, DEPRESSED)%>%
      dplyr::summarize(count = n())%>%dplyr::mutate(percentage = count/sum(count), pos = (cumsum(percentage) - 0.5 * percentage))
    #mapping strings to factors for income brackets for ordering
    df$scode<-mapvalues(df$ANNUAL_HH_INCOME, from = c("$0 - 4,999", "$5,000 - 9,999", "$10,000 - 14,999", "$15,000 - 19,999", "$20,000 - 24,999", "$25,000 - 34,999", "$35,000 - 44,999", "$45,000 - 54,999", "$55,000 - 64,999", "$65,000 - 74,999", "$75,000 - 99,999", "$100,000 +"), to=c("1", "2", "3", "4", "5", "6","7", "8", "9","10", "11", "12"))
    
      p<-ggplot(data= df, aes(x =ANNUAL_HH_INCOME, y=percentage, fill=DEPRESSED)) + geom_bar(stat="identity", position ="stack") + labs(title = "") + ggtitle("") + theme(legend.title =element_text(size=8), legend.text = element_text(size=6), plot.title=element_text(size=10, face="bold"), axis.title.x = element_text(size=7), axis.text=element_text(size=6))+ geom_text(data = subset(df,percentage>=0.05), aes(y=pos, label = paste(round(percentage*100,1), "%")), size=3, color="black")+ scale_x_discrete(limits= c("$0 - 4,999", "$5,000 - 9,999", "$10,000 - 14,999", "$15,000 - 19,999", "$20,000 - 24,999", "$25,000 - 34,999", "$35,000 - 44,999", "$45,000 - 54,999", "$55,000 - 64,999", "$65,000 - 74,999", "$75,000 - 99,999", "$100,000 +"), expand=c(0,-2)) + xlab("") + ylab("% of Total Count Depressed") + scale_y_continuous(expand=c(0,0)) + coord_flip() 
ggplotly(p)%>%layout(margin=list(b=150))
})
```

***
Bar Graph Interpretation:

The happiest divorced individuals are in the $75,000 - 99,000 income bracket with a never depressed rate of 90.91%. For respondents who are living with a partner, the happiests subjects are in the $65,000 to 74,999 income bracket with a never depressed rate of 90%. Widowed middle-class earners in the $45,000 to 54,999 income bracket are the happiest with a never depressed rate of 100%. Separated individuals in the $100,000 + bracket are the happiest with 100% of subjects never being depressed.The happiest married people are in the highest income bracket of $100,000 + with a never depressed rate of 87.10%. 

Notice that with the exception of widowers, the happiest respondents are also higher-income earners.


### Box Plot: Ratio Income Poverty v. Severe Depression by Gender
```{r, echo=FALSE}
df1<- read.csv("mentalHealth.reformatted.csv", stringsAsFactors = FALSE)
df2<- read.csv("demographics.reformatted.csv", stringsAsFactors = FALSE)
df<-merge(df1, df2)
df$scode<-mapvalues(df$ANNUAL_HH_INCOME, from = c("Nearly every day", "More than half the days", "Several days", "Never"), to=c("1", "2", "3", "4"))
selectInput(input = "gender",
              label = "Select a Gender: ",
              c("Male", "Female"), selected = "Male")
plotlyOutput("boxplot", width = "auto", height = "auto")
output$boxplot<-renderPlotly({
  df1<- read.csv("mentalHealth.reformatted.csv", stringsAsFactors = FALSE)
  df2<- read.csv("demographics.reformatted.csv", stringsAsFactors = FALSE)
  df<-merge(df1, df2)
      df <-df %>%
      dplyr::select(GENDER, RATIO_INCOME_POVERTY, BETTER_DEAD)%>%
      dplyr::filter(GENDER%in%c(input$gender), RATIO_INCOME_POVERTY!="null",BETTER_DEAD!="null")%>%
      dplyr::group_by(GENDER, RATIO_INCOME_POVERTY, BETTER_DEAD)
      p<-ggplot(data= df, aes(x =BETTER_DEAD, y=as.numeric(as.character(RATIO_INCOME_POVERTY)), fill=BETTER_DEAD)) + geom_boxplot() + xlim(c("Nearly every day", "More than half the days", "Several days", "Never")) + ylim(0,5) + xlab("Often Feels Better Dead") + ylab("Ratio Income Poverty") + theme(legend.position="none")
ggplotly(p)%>%layout(margin=list(b=150))
})
    

```

***
Boxplot Interpretation: 

For male respondents in this survey, those who feel severely depressed nearly everyday have a median income to poverty ratio of 0.63 while those who are never severely depressed have a median income to poverty ratio of 2.18. In general, for male respondents, as the severity of depression increases the income to poverty ratio decreases.

Female respondents who are depressed nearly every day and more than half the days have a median income to poverty ratio of 1 and 0.91 respectively. Females who feel severely depressed for several days have a median income to poverty ratio of 1.30. Notice that females who are never severely depressed have a median income to poverity ratio of 2.02.
